---
title: 'Week 9-10: Prepare - Heat maps, Spatial charts, and Contour charts'
author: "Shani Kumar"
output:
  word_document: default
  pdf_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises: Charts

You need to submit 3 heat maps, 3 spatial charts and 3 contour charts using Tableau or PowerBI, Python and R using the data below (or your own datasets). You can also use D3. You can choose which library to use in Python or R, documentation is provided to help you decide and as you start to play around in the libraries, you will decide which you prefer.

**Data source**
We are using dataset from [Data Source URL](https://content.bellevue.edu/cst/dsc/640/datasets/ex5-2.zip) file.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# import libraries
library(RColorBrewer)
library(maps)
library('ggplot2')
```
```{r read, echo=FALSE}
costco = read.csv2(file = "costcos-geocoded.csv", header = TRUE, sep=',', dec = '.')
ppg2008 = read.csv2(file = "ppg2008.csv", header = TRUE, sep=',', dec = '.')


# Show the data sample
head(costco)
head(ppg2008)

```

### Data structure:
```{r echo=FALSE}
str(costco)
str(ppg2008)
```

\pagebreak

### Construct Charts:

**Heat Map**

```{r echo=FALSE}
# Create a heatmap
heatmap(as.matrix(ppg2008[,2:ncol(ppg2008)]), Rowv = NA, Colv = NA, 
                         col = brewer.pal(9, "Blues"),
                        scale = "column", margins = c(5,10))
```

\pagebreak

**Spatial Plot**

```{r echo=FALSE}
map(database="state", col="#cccccc")
symbols(costco$Longitude, costco$Latitude, bg="#e2373f", fg="#ffffff",
  lwd=0.5, circles=rep(1, length(costco$Longitude)),
  inches=0.05, add=TRUE)
```

\pagebreak

**Countour plot**

```{r echo=FALSE}
ggplot(ppg2008, aes(x=FTM, y=FTA))+
  theme_bw()+
  geom_point(alpha=0.1, col='red')+
  geom_density2d(color='black')+
  ggtitle('Contour Plot')+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x='FTM', y='FTA')
```

