---
title: 'Week 11-12: Prepare - Histograms, Box Plots, & Bullet Charts'
author: "Shani Kumar"
output:
  html_document:
    df_print: paged
  pdf_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises: Charts

You need to submit 3 histograms, 3 box plots, 3 bullet charts, and 3 additional charts of your choice (can be an existing chart type we’ve already done, but it must be done in a new way or it can be an entirely new chart type) using Tableau or PowerBI, Python and R using the data below (or your own datasets). You can also use D3. You can choose which library to use in Python or R, documentation is provided to help you decide and as you start to play around in the libraries, you will decide which you prefer.

**Data source**
We are using dataset from [Data Source URL](https://content.bellevue.edu/cst/dsc/640/datasets/ex6-2.zip) file.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# import libraries
library(dplyr)
library(plotly)
```
```{r read, echo=FALSE}
birthrate <- read.csv("ex6-2/birth-rate.csv", header = TRUE, sep=',', dec = '.')
crimerate <- read.csv("ex6-2/crimeratesbystate-formatted.csv", header = TRUE, sep=',', dec = '.') 
education <- read.csv("ex6-2/education.csv", header = TRUE, sep=',', dec = '.') 

# Show the data sample
head(birthrate)

# format year columns
colnames(birthrate) <- gsub("X", "", colnames(birthrate))
head(crimerate)
head(education)
```

### Data structure:
```{r echo=FALSE}
str(birthrate)
str(crimerate)
str(education)
```

\pagebreak

### Construct Charts:

**Histogram**

```{r echo=FALSE}
# Create a Histogram
#options(repr.plot.width = 4, repr.plot.height = 4)

# create pivotted data for plotting
birthrate_hist <- reshape2::melt(birthrate, id=c("Country")) %>%
  dplyr::mutate("Country" = as.character(Country),
                "Year" = as.character(variable),
                "rate" = value) %>% 
  dplyr::select(c("Country","Year","rate"))

# create histogram of birth rate
hist(birthrate_hist$rate)

```

\pagebreak

**Box Plot**

```{r echo=FALSE}
# Create Box plot
birthrate_box <- birthrate_hist %>% 
  dplyr::filter(Country %in% c("United States"))

boxplot(birthrate_box$rate ~ birthrate_box$Country)

```

\pagebreak

**Bullet Chart**

```{r echo=FALSE}
# Create bullet chart
fig <- plot_ly(
    type = "indicator",
    mode = "number+gauge+delta",
    value = birthrate_box[1,3],
    domain = list(x = c(0, 1), y= c(0, 1)),
    title = list(text = "Birth Rate"),
    delta = list(reference = 30),
    gauge = list(
        shape = "bullet",
        axis = list(range = list(NULL, 30)),
        threshold = list(
            line = list(color = "red", width = 2),
            thickness = 0.75,
            value = 280),
        steps = list(
            list(range = c(0, 15), color = "lightgray"),
            list(range = c(15, 25), color = "gray"))),
    height = 150, width = 700) 
fig <- fig %>%
    layout(margin = list(l= 150, r= 10))

fig
```

```{r echo=FALSE}
ggplot(birthrate_box, aes(x = Year, y = rate)) 
+ geom_line() 

```

